Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2021-07-22 edubots_surveys.Rnw
Completed first draft of paper
Closes #47 (by @fbarranco before
Closes #24 (by @fbarranco and @oliatsi)
Closes #22 (by @fbarranco and @amorag)
Closes #19 (done before)
Closes #17 (not really done, finally eliminated)
Closes #16 Done, discusses, added conclusions
Closes #15 Done, also added results and conclusions

% Add something about intro
% RQ1 discussion about messaging app: they consider a IM the Moodle messaging
% Olia: it's not so easy to change habits, even in the case of major crisis. - JJ
% Relationship between changes, gender and discipline (Olia) - JJ
% 
Our initial intention in the design of these surveys was to probe the opinions of tertiary education teachers in the introduction of chatbot technologies in class. We wanted to find the answer to four research questions, of which the first was to check whether teachers already used some kind of messaging application, and the one they preferred. Depending on the answer, different technologies could be (or not) available for the introduction of chatbots. In this case, the design of the survey introduced some ambiguity on what we really consider ``messaging technology'', that was eventually resolved by the answers; apparently, anything that sends messages (even if they are not instant, even if they do not have a dedicated application) was the concept in the mind of the surveyed.

@@

% Rq4: they seem to prefer interaction with students - Olia 
% Alberto: using new metric so grade students, and also abotu horizontality

@@

% Noorhan suggest that using some simple things could increase the perceived usefulness - JJ

% What kidn of chatbots are preferred? (suggested by Fran)

% No remarkable differences between different groups.

% Only some COVID-response relationship.

% Future study (suggested by Noorhan): how COVID influenced adoption of these technologies.

% Emotinal, mood for future work... Chatbots perform sentiment analysis

%This section is repeated
%\section*{Acknowledgements}

@@

The datasets generated and analysed during the current study are available in the [NAME] repository, accesible from [PERSISTENT WEB LINK TO DATASETS]

>>>>

Our initial intention in the design of these surveys was to probe the opinions of tertiary education teachers in the introduction of chatbot technologies in class. We wanted to find the answer to four research questions, of which the first was to check whether teachers already used some kind of messaging application, and the one they preferred. Depending on the answer, different technologies could be (or not) available for the introduction of chatbots. In this case, the design of the survey introduced some ambiguity on what we really consider ``messaging technology'', that was eventually resolved by the answers; apparently, anything that sends messages (even if they are not instant, even if they do not have a dedicated application) was the concept in the mind of the surveyed.

@@

@@

% Noorhan suggest that using some simple things could increase the perceived usefulness - JJ
As in many other methods of technology acceptance, answers to the survey suggest that the introduction of simple, and institution-supported, instant messaging and chatbot technologies would increase the perceived usefulness (which is one of the key metrics in technology acceptance models). Since most institutions already have some {\em virtual campus} or learning management system, adding some easy automation, or connection to personal instant messaging tools, could really help onboard the learning community on these new technologies.

% What kidn of chatbots are preferred? (suggested by Fran)
Once that initial introduction is done, teachers (and students as well) will probably prefer the kind of bots that alleviate bureaucratic or repetitive tasks, such as answering frequent questions or answers on class or assignment schedules. These will help the introduction of more complex chatbots that will affect more directly learning outcomes, such as chatbots that help students integrate in the class, or are able to identify (and address) learning problems in students or in groups of them. These as well should be accompanied by analytics on student interaction, as well as possibly some natural language processing (in vernacular language) that will help assess the general mood of the class, and how different material (or external factors) affect it.

% No remarkable differences between different groups.
This introduction of chatbots and chatbot technology will be helped by the fact that no discernible differences are found between different groups. Even if chatbots do have some potential for personalization or customization, based on the class material, student and teaching staff, the general introduction of the technology can be done in a general way and for all disciplines, teaching experience and gender.

% Only some COVID-response relationship.
In our survey, we couldn't find an answer to the right moment for doing this introduction of chatbot technologies. Response to questions related to change of behavior during and after the pandemic do not suggest that major (or minor) crisis could be an opportunity to introduce new technologies, since it does not bring major changes in attitude. A minor crisis would be, for instance, rollout of a new higher education law\footnote{We seem to have one of these, at least in Spain, every 10 years or so}, or introduction of new degrees. Although external changes do offer the chance of piloting new technologies, they do not seem to bring changes in attitudes in the teaching staff (which, after all, is bound to be the same). In absence of a clear answer in this direction, the right moment to introduce new technologies is always when the IT and managing staff is ready to support it (since, as we have seen before, "official" support is one of the factors that is most valued by teachers).

At any rate, the general feeling that transpires from the survey is that it is essential for any institution to take into account the stakeholder's opinion when introducing chatbots. This is true almost across the board, but in the case of chatbots (and instant messaging applications) their inmediacy and the fact that they can enter into what we could call the private sphere makes this even more necessary.

% Future study (suggested by Noorhan): how COVID influenced adoption of these technologies.
There are several future lines of work informed by our results of this survey. For instance, the rollout of extensive videoconferencing and virtual teaching solutions that the COVID pandemic has brought has also taught us a series of lessons. It increases isolation, for instance, and decreases the amount of synchronous contact that happens in the fringes of the classroom: teaching staff offices, before and after class. A future line of work would be focused on exactly this, and what kind of needs could be covered by chatbot technology. We will create a series of international surveys that will investigate this.

% Emotinal, mood for future work... Chatbots perform sentiment analysis
Finally, the full extent of chatbot technology is not really being examined in these surveys. They can be connected to natural language processing engines with sentiment analysis as well as other analytics. Examining the mood of the class, as response to new material, assignment, exams or external events might help will help any student-centered teaching strategy, which will help students (and teachers) reach their learning objectives. This line of work, however, will be based on the introduction of some pilot study, combined with initial opinion assesment on the students.

@@

The datasets generated and analysed during the current study are available in the {\tt edubots-paper} repository, accesible from {\tt https://github.com/JJ/edubots-paper/tree/main/data}.
